---
title: "Controlling the sketch"
description: >
  The shared sketch parameters — roughness, bowing, n_passes, and seed — and how
  each changes the hand-drawn look.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 4, dpi = 120, dev = "ragg_png",
  out.width = "100%", fig.align = "center"
)
library(ggplot2)
library(ggsketch)
options(ggsketch.seed = 1)
```

Every sketch geom shares four parameters that shape the hand-drawn look:
`roughness`, `bowing`, `n_passes`, and `seed`. They are layer parameters (not
aesthetics), so you pass them to the `geom_sketch_*()` call.

## `roughness`

How far points are displaced. `0` is ruler-straight; the default is `1`; values
above `~3` get loose and scribbly.

```{r roughness}
x  <- seq(0, 10, length.out = 50)
d  <- data.frame(x = x, y = sin(x))

ggplot(d, aes(x, y)) +
  geom_sketch_line(aes(colour = "0.0  (straight)"), roughness = 0,   seed = 2) +
  geom_sketch_line(aes(colour = "1.0  (default)"),  roughness = 1,   seed = 2) +
  geom_sketch_line(aes(colour = "2.5"),             roughness = 2.5, seed = 2) +
  geom_sketch_line(aes(colour = "5.0  (loose)"),    roughness = 5,   seed = 2) +
  scale_colour_brewer("roughness", palette = "Set1") +
  labs(title = "The roughness dial") +
  theme_sketch()
```

On bars:

```{r roughness-bars}
df <- data.frame(x = c("A", "B", "C"), y = c(4, 6, 3))
for (r in c(0, 1, 3)) {
  print(
    ggplot(df, aes(x, y)) +
      geom_sketch_col(fill = "#7BAFD4", roughness = r, seed = 1L) +
      labs(subtitle = paste("roughness =", r), x = NULL) +
      theme_sketch(base_size = 9)
  )
}
```

## `bowing`

How much line segments bow (curve) between their endpoints. Higher values give a
springier, more relaxed stroke.

```{r bowing}
seg <- data.frame(x = 1, y = 1, xend = 6, yend = 4)
for (b in c(0, 2, 5)) {
  print(
    ggplot(seg) +
      geom_sketch_segment(aes(x = x, y = y, xend = xend, yend = yend),
                          bowing = b, roughness = 0.6, linewidth = 1,
                          seed = 7L) +
      labs(subtitle = paste("bowing =", b)) +
      theme_sketch(base_size = 9)
  )
}
```

## `n_passes`

How many overlaid strokes to draw. `2` (the default) gives the classic
"double-stroke" sketch look; `1` is a single pass; `3+` looks heavily worked.

```{r passes}
for (n in c(1, 2, 4)) {
  print(
    ggplot(d, aes(x, y)) +
      geom_sketch_line(n_passes = n, roughness = 1.5, linewidth = 0.7,
                       colour = "grey20", seed = 3L) +
      labs(subtitle = paste("n_passes =", n)) +
      theme_sketch(base_size = 9)
  )
}
```

## `medium` (line geoms)

The line family — `geom_sketch_line()`, `geom_sketch_path()`,
`geom_sketch_segment()`, `geom_sketch_step()` — also takes a `medium`, which sets
*how* the stroke is laid down: a tapering pencil, a wet ink line, a crisp
fountain pen, a thin ballpoint, a broad brush, charcoal, soft pastel, dusty
chalk (best on `paper = "chalkboard"`), marker, a translucent highlighter
band, or crayon. `sketch_media()` lists them.

```{r medium}
for (m in c("pencil", "ink", "brush", "charcoal")) {
  print(
    ggplot(d, aes(x, y)) +
      geom_sketch_line(medium = m, linewidth = 1.2, seed = 2L) +
      labs(subtitle = paste("medium =", m)) +
      theme_sketch(base_size = 9)
  )
}
```

You can also *map* `medium` to a variable with `scale_medium_discrete()` (see the
gallery). The default, `"pen"`, reproduces the classic constant-width sketch line,
so existing plots are unchanged.

## `seed` — reproducible wobble

The sketch is random, but **seeded**. A given `seed` always produces the same
drawing, so figures are reproducible across renders and devices. Change the seed
for a fresh "hand".

```{r seed}
d2 <- data.frame(x = 1:8, y = c(2, 5, 3, 8, 6, 9, 7, 10))
for (s in c(1, 1, 42)) {
  print(
    ggplot(d2, aes(x, y)) +
      geom_sketch_line(linewidth = 1, seed = s) +
      geom_sketch_point(size = 3, seed = s + 100) +
      labs(subtitle = paste("seed =", s)) +
      theme_sketch(base_size = 9)
  )
}
```

The first two panels are identical (same seed); the third differs.

### Session-wide default

If you do not pass `seed`, geoms use `getOption("ggsketch.seed", 1L)`. Set it
once to make a whole document reproducible:

```{r option, eval = FALSE}
options(ggsketch.seed = 1L)
```

ggsketch never touches your global random state: each randomized routine draws
from its own seeded stream and restores `.Random.seed` afterward, so it will not
interfere with `set.seed()` elsewhere in your analysis.
