---
title: "Gallery — every geom"
description: >
  A tour of every ggsketch geom with runnable examples, from bars and lines to
  histograms, violins, intervals, Q-Q plots, quantile fits, reference lines,
  curves, and annotations — plus the 2.0 additions: dumbbell, slope, beeswarm,
  ridgeline, streamgraph, waffle, treemap and calendar charts, hand-drawn
  networks and sf maps, drawing media, watercolour fills, textured paper, and
  hand-drawn coordinate frames.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  message    = FALSE,
  warning    = FALSE,
  fig.width  = 7,
  fig.height = 4.2,
  dpi        = 120,
  dev        = "ragg_png",
  out.width  = "100%",
  fig.align  = "center"
)
library(ggplot2)
library(ggsketch)
set.seed(1)
options(ggsketch.seed = 1)
# Use a handwriting font for *all* plot text (titles, axes, legends), not just
# the labels drawn by geom_sketch_text()/geom_sketch_bracket(). This sets the
# default base_family for every theme_sketch() call below.
options(ggsketch.base_family = "auto")
```

Everything here is a real ggplot2 layer, so it composes with `aes()`, stats,
scales, facets, and coords — and renders on any device. Every example sets a
`seed` so the wobble is reproducible.

Every plot uses a handwriting font throughout via
`options(ggsketch.base_family = "auto")` in setup; without it, `theme_sketch()`
keeps the device default font and only the labels that geoms draw themselves
(`geom_sketch_text()`, `geom_sketch_bracket()`) are hand-drawn.

## Bars and columns

`geom_sketch_col()` draws a roughened outline with a hachure (pencil-shading)
fill.

```{r col-basic}
sales <- data.frame(product = c("Alpha", "Bravo", "Charlie", "Delta", "Echo"),
                    units   = c(34, 51, 22, 47, 39))

ggplot(sales, aes(product, units)) +
  geom_sketch_col(fill = "#7BAFD4", seed = 1L) +
  labs(title = "Units sold", x = NULL) +
  theme_sketch()
```

Map `fill` to a variable like any ggplot2 bar. Each bar gets its own seed offset,
so no two bars wobble identically.

```{r col-mapped}
ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(seed = 2L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Mapped fill", x = NULL) +
  theme_sketch()
```

`geom_sketch_bar()` counts rows for you (like `geom_bar()`):

```{r bar-count}
ggplot(mpg, aes(class)) +
  geom_sketch_bar(fill = "#C39BD3", seed = 3L) +
  labs(title = "Vehicle count by class", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Bars flip and stack just like the originals:

```{r col-flip}
ggplot(sales, aes(reorder(product, units), units)) +
  geom_sketch_col(fill = "#F1948A", seed = 1L) +
  coord_flip() +
  labs(title = "Horizontal bars", x = NULL) +
  theme_sketch()
```

## Lollipops

`geom_sketch_lollipop()` is a tidy alternative to bars for ranked or sparse
values: a roughened stem from a `baseline` capped with a sketch point. The
value axis expands to include the baseline.

```{r lollipop}
ggplot(sales, aes(reorder(product, units), units)) +
  geom_sketch_lollipop(colour = "#7B241C", seed = 1L) +
  labs(title = "Units sold", x = NULL) +
  theme_sketch()
```

## Histograms and frequency polygons

`geom_sketch_histogram()` bins a continuous variable and draws hand-drawn bars;
`geom_sketch_freqpoly()` draws the same counts as a roughened line.

```{r histogram}
ggplot(faithful, aes(eruptions)) +
  geom_sketch_histogram(fill = "#7BAFD4", bins = 20, seed = 1L) +
  labs(title = "Old Faithful eruption times") +
  theme_sketch()
```

```{r freqpoly}
ggplot(mpg, aes(hwy, colour = drv)) +
  geom_sketch_freqpoly(bins = 15, linewidth = 0.9, seed = 2L) +
  labs(title = "Highway mpg by drivetrain", x = "hwy") +
  theme_sketch()
```

## Dot plots

`geom_sketch_dotplot()` bins the data and stacks one roughened circular dot per
observation. The dots are sized by the bin width, so the count axis is
approximate — turn it off for a clean look.

```{r dotplot}
ggplot(faithful, aes(eruptions)) +
  geom_sketch_dotplot(binwidth = 0.12, fill = "#7BAFD4", seed = 1L) +
  scale_y_continuous(NULL, breaks = NULL) +
  labs(title = "Old Faithful eruption times") +
  theme_sketch()
```

## Empirical CDF

`geom_sketch_ecdf()` draws the empirical cumulative distribution as a hand-drawn
stairstep.

```{r ecdf}
ggplot(mpg, aes(hwy, colour = drv)) +
  geom_sketch_ecdf(linewidth = 0.8, seed = 1L) +
  labs(title = "Highway mpg ECDF by drivetrain", y = "F(x)") +
  theme_sketch()
```

## Lines, paths, and points

```{r line}
econ <- economics[economics$date > as.Date("2000-01-01"), ]
ggplot(econ, aes(date, unemploy)) +
  geom_sketch_line(colour = "steelblue", linewidth = 0.8, seed = 1L) +
  labs(title = "US unemployment", x = NULL, y = "thousands") +
  theme_sketch()
```

`geom_sketch_point()` draws each point as a small roughened ellipse:

```{r point}
ggplot(mtcars, aes(wt, mpg, colour = factor(cyl))) +
  geom_sketch_point(size = 4, seed = 1L) +
  scale_colour_brewer("cylinders", palette = "Dark2") +
  labs(title = "Fuel economy vs weight") +
  theme_sketch()
```

Lines and points compose like any layers:

```{r line-point}
df <- data.frame(x = 1:12, y = c(3, 5, 4, 7, 6, 9, 8, 11, 9, 12, 11, 14))
ggplot(df, aes(x, y)) +
  geom_sketch_line(colour = "grey40", seed = 3L) +
  geom_sketch_point(size = 4, colour = "firebrick", seed = 8L) +
  labs(title = "Trend with markers") +
  theme_sketch()
```

Multiple groups, one seed per group:

```{r line-groups}
ggplot(ggplot2::economics_long, aes(date, value01, colour = variable)) +
  geom_sketch_line(seed = 5L) +
  labs(title = "Five series, hand-drawn", x = NULL, y = NULL) +
  theme_sketch()
```

## Point sizes

`size` behaves like any ggplot2 point size. Set it to a constant for bigger or
smaller markers:

```{r point-size-fixed}
sz <- data.frame(x = 1:5, y = 1, s = c(2, 4, 6, 9, 13))
ggplot(sz, aes(x, y)) +
  geom_sketch_point(size = sz$s, colour = "#2E86C1", seed = 1L) +
  labs(title = "Fixed point sizes (2 → 13)", x = NULL, y = NULL) +
  theme_sketch()
```

Or map `size` to a variable for a bubble chart — pair it with
`scale_size_area()` so the *area* (not radius) encodes the value:

```{r point-size-mapped}
ggplot(mtcars, aes(wt, mpg, size = hp, colour = factor(cyl))) +
  geom_sketch_point(alpha = 0.9, seed = 2L) +
  scale_size_area("horsepower", max_size = 14) +
  scale_colour_brewer("cylinders", palette = "Dark2") +
  labs(title = "Bubble chart: size = horsepower") +
  theme_sketch()
```

A small-multiples sweep of a single size aesthetic:

```{r point-size-grid}
grid <- expand.grid(x = 1:6, y = 1:3)
grid$s <- seq(1.5, 11, length.out = nrow(grid))
ggplot(grid, aes(x, y, size = s)) +
  geom_sketch_point(colour = "#884EA0", show.legend = FALSE, seed = 3L) +
  scale_size_identity() +
  labs(title = "Increasing point size", x = NULL, y = NULL) +
  theme_sketch()
```

## Point roughness

For `geom_sketch_point()`, `roughness` is a *mappable aesthetic*. As a constant
it sets how wobbly every marker is — from clean circles up to very shaky:

```{r point-rough-const}
rg <- data.frame(x = 1:4, y = 1, r = c(0, 0.4, 0.9, 1.6))
ggplot(rg, aes(x, y)) +
  geom_sketch_point(aes(roughness = I(r)), size = 14, colour = "#2E86C1",
                    seed = 1L) +
  geom_sketch_text(aes(label = r), nudge_y = -0.5, size = 6) +
  labs(title = "roughness 0 → 1.6 (constant per point)",
       x = NULL, y = NULL) +
  ylim(0.3, 1.3) +
  theme_sketch() +
  theme(axis.text = element_blank())
```

Map it to a variable and the values are rescaled to a legible band by
`scale_roughness_continuous()` (applied automatically, default `c(0.01, 0.75)`),
so points can encode a third variable through how shaky they look:

```{r point-rough-mapped}
ggplot(mtcars, aes(wt, mpg, roughness = hp, colour = factor(cyl))) +
  geom_sketch_point(size = 4, seed = 1L) +
  scale_colour_brewer("cylinders", palette = "Dark2") +
  labs(title = "roughness mapped to horsepower") +
  theme_sketch()
```

Use `I()` to pass raw roughness through unscaled, or
`scale_roughness_continuous(range = ...)` to widen the band.

## Jitter and count

`geom_sketch_jitter()` spreads overplotted points; `geom_sketch_count()` sizes a
single point by how many observations sit there.

```{r jitter}
ggplot(mpg, aes(class, hwy)) +
  geom_sketch_jitter(width = 0.2, height = 0, colour = "#5D6D7E",
                     size = 2, seed = 1L) +
  labs(title = "Jittered highway mpg", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r count}
ggplot(mpg, aes(cty, hwy)) +
  geom_sketch_count(colour = "#C0392B", seed = 2L) +
  scale_size_area(max_size = 8) +
  labs(title = "Overplot count") +
  theme_sketch()
```

## Rectangles and tiles

```{r rect}
rects <- data.frame(xmin = c(1, 3, 5), xmax = c(2, 4, 6),
                    ymin = 0, ymax = c(2, 4, 3))
ggplot(rects) +
  geom_sketch_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                       fill = factor(xmin)),
                   seed = 1L, show.legend = FALSE) +
  labs(title = "geom_sketch_rect()") +
  theme_sketch()
```

A sketchy heatmap with `geom_sketch_tile()`:

```{r tile}
td <- expand.grid(x = 1:8, y = 1:6)
td$z <- td$x + td$y
ggplot(td, aes(x, y, fill = z)) +
  geom_sketch_tile(seed = 2L, hachure_gap = 0.18) +
  scale_fill_viridis_c() +
  labs(title = "geom_sketch_tile()") +
  theme_sketch()
```

A sketchy 2-D bin heatmap with `geom_sketch_bin2d()` (cells default to a hachure
fill, shaded by count):

```{r bin2d}
ggplot(faithful, aes(eruptions, waiting)) +
  geom_sketch_bin2d(bins = 12, seed = 3L) +
  scale_fill_viridis_c() +
  labs(title = "geom_sketch_bin2d()") +
  theme_sketch()
```

`geom_sketch_hex()` bins into hexagons instead (needs the **hexbin** package):

```{r hex, eval = requireNamespace("hexbin", quietly = TRUE)}
ggplot(faithful, aes(eruptions, waiting)) +
  geom_sketch_hex(bins = 12, seed = 4L) +
  scale_fill_viridis_c(option = "magma") +
  labs(title = "geom_sketch_hex()") +
  theme_sketch()
```

## Polygons, ribbons, areas, and densities

Concave polygons fill correctly (the hachure respects every notch):

```{r polygon}
ang  <- seq(0, 2 * pi, length.out = 11)[-11]
r    <- rep(c(1, 0.45), length.out = 10)
star <- data.frame(x = r * cos(ang), y = r * sin(ang))
ggplot(star, aes(x, y)) +
  geom_sketch_polygon(fill = "tomato", seed = 1L) +
  coord_equal() +
  labs(title = "A concave star") +
  theme_sketch()
```

```{r ribbon-area}
band <- data.frame(x = 1:20)
band$y  <- 10 + 5 * sin(seq(0, 3 * pi, length.out = 20))
band$lo <- band$y - 2
band$hi <- band$y + 2

ggplot(band, aes(x)) +
  geom_sketch_ribbon(aes(ymin = lo, ymax = hi), fill = "plum", seed = 2L) +
  geom_sketch_line(aes(y = y), seed = 3L) +
  labs(title = "Ribbon + line") +
  theme_sketch()
```

```{r area}
ggplot(band, aes(x, y)) +
  geom_sketch_area(fill = "lightgreen", seed = 3L) +
  labs(title = "geom_sketch_area()") +
  theme_sketch()
```

```{r density}
ggplot(faithful, aes(eruptions)) +
  geom_sketch_density(fill = "khaki", seed = 4L) +
  labs(title = "Old Faithful eruptions") +
  theme_sketch()
```

## Violins

`geom_sketch_violin()` mirrors a kernel density into a closed polygon and
hachure-fills it.

```{r violin}
ggplot(mpg, aes(class, hwy, fill = class)) +
  geom_sketch_violin(seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "Highway mpg distribution by class", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Smooths

A hand-drawn fit with a roughened confidence band:

```{r smooth}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(seed = 1L) +
  geom_sketch_smooth(method = "lm", formula = y ~ x, seed = 2L) +
  labs(title = "Linear fit with CI band") +
  theme_sketch()
```

```{r smooth-loess}
ggplot(mpg, aes(displ, hwy)) +
  geom_sketch_point(colour = "grey50", seed = 1L) +
  geom_sketch_smooth(seed = 2L, colour = "darkorange") +
  labs(title = "loess fit") +
  theme_sketch()
```

## Function curves

`geom_sketch_function()` sketches an analytic curve over the x range, for example
to overlay a theoretical density.

```{r function}
ggplot(data.frame(x = c(-4, 4)), aes(x)) +
  geom_sketch_function(fun = dnorm, colour = "#2E86C1", linewidth = 0.9,
                       seed = 1L) +
  geom_sketch_function(fun = dnorm, args = list(sd = 1.6),
                       colour = "#C0392B", linewidth = 0.9, seed = 2L) +
  labs(title = "Two normal densities", y = "density") +
  theme_sketch()
```

## Q-Q plots

`geom_sketch_qq()` draws the quantile-quantile points and `geom_sketch_qq_line()`
the reference line. Map data to the `sample` aesthetic.

```{r qq}
ggplot(mtcars, aes(sample = mpg)) +
  geom_sketch_qq(size = 2.5, seed = 1L) +
  geom_sketch_qq_line(colour = "#C8553D", linewidth = 0.8, seed = 2L) +
  labs(title = "Normal Q-Q plot of mpg", x = "theoretical", y = "sample") +
  theme_sketch()
```

## Quantile regression

`geom_sketch_quantile()` fits and draws quantile regression lines (requires the
optional **quantreg** package).

```{r quantile, eval = requireNamespace("quantreg", quietly = TRUE)}
ggplot(mpg, aes(displ, hwy)) +
  geom_sketch_point(colour = "grey60", seed = 1L) +
  geom_sketch_quantile(quantiles = c(0.1, 0.5, 0.9), colour = "#6C3483",
                       linewidth = 0.9, seed = 2L) +
  labs(title = "10th / 50th / 90th percentile fits") +
  theme_sketch()
```

## Circles and ellipses

Radii are in data units, so use `coord_equal()` for true circles:

```{r circle}
cdf <- data.frame(x = c(1, 3, 2), y = c(1, 1, 2.5),
                  r = c(0.6, 0.9, 0.5), grp = c("a", "b", "c"))
ggplot(cdf, aes(x, y, r = r, fill = grp)) +
  geom_sketch_circle(seed = 1L, show.legend = FALSE) +
  coord_equal() +
  labs(title = "geom_sketch_circle()") +
  theme_sketch()
```

```{r ellipse}
edf <- data.frame(x = c(1, 3), y = c(1, 2), a = c(1.4, 0.8), b = c(0.6, 1.2))
ggplot(edf, aes(x, y, a = a, b = b, fill = factor(x))) +
  geom_sketch_ellipse(seed = 2L, show.legend = FALSE) +
  coord_equal() +
  labs(title = "geom_sketch_ellipse()") +
  theme_sketch()
```

## Segments and steps

```{r segment}
sdf <- data.frame(x = 1:4, y = c(1, 3, 2, 4),
                  xend = 2:5, yend = c(3, 1, 4, 2))
ggplot(sdf) +
  geom_sketch_segment(aes(x = x, y = y, xend = xend, yend = yend),
                      colour = "darkgreen", linewidth = 1, seed = 3L) +
  labs(title = "geom_sketch_segment()") +
  theme_sketch()
```

```{r step}
stp <- data.frame(x = 1:8, y = c(1, 3, 2, 5, 4, 6, 5, 8))
ggplot(stp, aes(x, y)) +
  geom_sketch_step(colour = "purple", linewidth = 1, seed = 4L) +
  geom_sketch_point(seed = 5L) +
  labs(title = "geom_sketch_step()") +
  theme_sketch()
```

## Curves and spokes

`geom_sketch_curve()` is a curved connector (a quadratic Bézier); `curvature`
sets how much it bends.

```{r curve}
cdf <- data.frame(x = c(1, 1, 1), y = c(1, 2, 3),
                  xend = c(4, 4, 4), yend = c(1, 2, 3))
ggplot(cdf, aes(x, y)) +
  geom_sketch_curve(aes(xend = xend, yend = yend), curvature = 0.4,
                    colour = "#1A5276", linewidth = 0.9, seed = 1L) +
  geom_sketch_point(seed = 2L) +
  geom_sketch_point(aes(x = xend, y = yend), seed = 3L) +
  labs(title = "geom_sketch_curve()") +
  theme_sketch()
```

`geom_sketch_spoke()` draws a segment from each point by `angle` and `radius` —
useful for vector fields.

```{r spoke}
field <- expand.grid(x = 1:6, y = 1:6)
field$angle  <- with(field, atan2(y - 3.5, x - 3.5))
field$radius <- 0.6
ggplot(field, aes(x, y)) +
  geom_sketch_spoke(aes(angle = angle, radius = radius),
                    colour = "#117A65", seed = 1L) +
  geom_sketch_point(size = 1.5, seed = 2L) +
  coord_equal() +
  labs(title = "geom_sketch_spoke()") +
  theme_sketch()
```

## Rugs

`geom_sketch_rug()` adds marginal ticks along the panel edges (`sides`).

```{r rug}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(colour = "#2C3E50", seed = 1L) +
  geom_sketch_rug(sides = "bl", colour = "#7B241C", seed = 2L) +
  labs(title = "Scatter with marginal rug") +
  theme_sketch()
```

## Intervals and uncertainty

The interval family draws hand-drawn ranges: `geom_sketch_linerange()`,
`geom_sketch_pointrange()`, `geom_sketch_errorbar()`, and
`geom_sketch_crossbar()`.

```{r intervals}
est <- data.frame(
  group = c("A", "B", "C", "D"),
  mean  = c(4.1, 5.6, 3.2, 6.0),
  lo    = c(3.2, 4.9, 2.4, 5.1),
  hi    = c(5.0, 6.4, 4.1, 6.8)
)
ggplot(est, aes(group, mean)) +
  geom_sketch_pointrange(aes(ymin = lo, ymax = hi), colour = "#1F618D",
                         seed = 1L) +
  labs(title = "Point estimates with 95% intervals", x = NULL) +
  theme_sketch()
```

```{r errorbar}
ggplot(est, aes(group, mean)) +
  geom_sketch_col(fill = "#AED6F1", width = 0.6, seed = 1L) +
  geom_sketch_errorbar(aes(ymin = lo, ymax = hi), width = 0.3, seed = 2L) +
  labs(title = "Bars with error bars", x = NULL) +
  theme_sketch()
```

```{r crossbar}
ggplot(est, aes(group, mean)) +
  geom_sketch_crossbar(aes(ymin = lo, ymax = hi), fill = "#FCF3CF",
                       fill_style = "hachure", seed = 3L) +
  labs(title = "geom_sketch_crossbar()", x = NULL) +
  theme_sketch()
```

## Reference lines

`geom_sketch_abline()`, `geom_sketch_hline()`, and `geom_sketch_vline()` span the
panel with a gentle wobble.

```{r reference}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(seed = 1L) +
  geom_sketch_hline(yintercept = 20, colour = "#C0392B", seed = 2L) +
  geom_sketch_vline(xintercept = 3.3, colour = "#2471A3", seed = 3L) +
  geom_sketch_abline(slope = -5, intercept = 37, colour = "#117864",
                     linetype = 2, seed = 4L) +
  labs(title = "Reference lines") +
  theme_sketch()
```

## Contours and 2-D density

`geom_sketch_contour()` draws contour lines of a surface (needs `z`);
`geom_sketch_density2d()` contours a 2-D kernel density estimate.

```{r contour}
ggplot(faithfuld, aes(waiting, eruptions, z = density)) +
  geom_sketch_contour(colour = "#2E4053", seed = 1L) +
  labs(title = "geom_sketch_contour()") +
  theme_sketch()
```

```{r density2d}
ggplot(faithful, aes(eruptions, waiting)) +
  geom_sketch_point(colour = "grey70", seed = 1L) +
  geom_sketch_density2d(colour = "#884EA0", linewidth = 0.7, seed = 2L) +
  labs(title = "geom_sketch_density2d()") +
  theme_sketch()
```

`geom_sketch_contour_filled()` fills the *bands* between levels instead of just
the lines. Each band may contain holes (the next level up, cut out); a
hole-aware filler keeps them empty, so `fill_style = "hachure"` works too.

```{r contour-filled}
ggplot(faithfuld, aes(waiting, eruptions, z = density)) +
  geom_sketch_contour_filled(seed = 1L) +
  labs(title = "geom_sketch_contour_filled()") +
  theme_sketch()
```

```{r density-2d-filled}
ggplot(faithful, aes(eruptions, waiting)) +
  geom_sketch_density_2d_filled(fill_style = "hachure", seed = 2L) +
  labs(title = "geom_sketch_density_2d_filled() (hachure)") +
  theme_sketch()
```

## Engraving and tonal shading

`geom_sketch_engrave()` shades a surface the way an etcher or banknote engraver
does: continuous tone is built from the *density* of hatch lines, with
cross-hatching deepening the shadows. It takes an `x`/`y`/`z` grid; high `z` is
dark. Unlike the fill-pattern packages it *computes* tone from geometry rather
than tiling a motif.

```{r engrave}
ggplot(faithfuld, aes(waiting, eruptions, z = density)) +
  geom_sketch_engrave(seed = 1L) +
  labs(title = "geom_sketch_engrave()") +
  theme_sketch()
```

With no pre-computed grid, shade raw points through a density stat:

```{r engrave-raw}
ggplot(faithful, aes(eruptions, waiting)) +
  geom_sketch_engrave(stat = "density_2d", contour = FALSE,
                      aes(z = after_stat(density)), seed = 1L) +
  labs(title = "geom_sketch_engrave() from raw points") +
  theme_sketch()
```

`geom_sketch_shade()` shades each polygon region with a *uniform* density set by
a `tone` aesthetic in `[0, 1]`, so a mapped value reads directly as darkness.
Mapping a raw variable to `tone` rescales it with `scale_tone_continuous()`
(alias `scale_engrave()`), just as `scale_size()` rescales size.

```{r shade}
hex <- data.frame(
  x = cos(seq(0, 2 * pi, length.out = 7))[-7],
  y = sin(seq(0, 2 * pi, length.out = 7))[-7]
)
regions <- do.call(rbind, lapply(1:3, function(k) {
  transform(hex, x = x + (k - 1) * 2.3, g = k, val = c(0.25, 0.55, 0.9)[k])
}))
ggplot(regions, aes(x, y, group = g)) +
  geom_sketch_shade(aes(tone = val), seed = 2L) +
  coord_equal() +
  labs(title = "geom_sketch_shade(): value → density") +
  theme_sketch()
```

## Text

The sketch of text is a *handwriting font*, not roughened glyphs.
`geom_sketch_text()` uses the first installed handwriting face (and falls back to
the device default otherwise).

```{r text}
lab <- data.frame(x = c(2, 4, 3), y = c(3, 4, 1.5),
                  txt = c("hand", "drawn", "labels"))
ggplot(lab, aes(x, y, label = txt)) +
  geom_sketch_point(size = 3, colour = "#C0392B") +
  geom_sketch_text(size = 7, nudge_y = 0.4) +
  labs(title = "geom_sketch_text()") +
  theme_sketch()
```

## Boxplots

A composed geom: rough IQR box, thick median, whiskers, and sketchy outliers.

```{r boxplot}
ggplot(mpg, aes(class, hwy)) +
  geom_sketch_boxplot(seed = 1L) +
  labs(title = "Highway mpg by class", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

By default the box is outline-only (its `fill` is `NA`). Give it a `fill` for a
solid box, or map `fill` and switch on a fill style for coloured, shaded boxes:

```{r boxplot-hachure}
ggplot(mpg, aes(class, hwy, fill = class)) +
  geom_sketch_boxplot(fill_style = "hachure", seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(title = "Hachure-filled boxes", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Annotations

`annotate_sketch()` adds one-off hand-drawn marks (no `aes()` inheritance):

```{r annotate}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(seed = 1L) +
  annotate_sketch("rect", xmin = 3, xmax = 4, ymin = 15, ymax = 22,
                  fill = NA, colour = "red", seed = 2L) +
  annotate_sketch("segment", x = 2, y = 32, xend = 3.4, yend = 21,
                  colour = "blue", linewidth = 1, seed = 3L) +
  annotate_sketch("circle", x = 5, y = 30, r = 0.4,
                  colour = "darkgreen", fill = "green", seed = 4L) +
  labs(title = "Highlighting with annotate_sketch()") +
  theme_sketch()
```

## Significance brackets

`geom_sketch_bracket()` draws a hand-drawn comparison bracket with an optional
handwriting label, for marking pairwise comparisons (a sketchy `ggsignif`).

```{r bracket}
brackets <- data.frame(
  xmin  = c(1, 2),
  xmax  = c(2, 3),
  y     = c(40, 45),
  label = c("p = 0.03", "n.s.")
)
ggplot(mpg, aes(drv, hwy)) +
  geom_sketch_boxplot(seed = 1L) +
  geom_sketch_bracket(
    data = brackets,
    aes(xmin = xmin, xmax = xmax, y = y, label = label),
    seed = 2L
  ) +
  labs(title = "Pairwise comparisons", x = "drivetrain") +
  theme_sketch()
```

## Pie & annotation toolkit

### Pie and donut charts

`geom_sketch_pie()` draws a hand-drawn pie sized by the `amount` aesthetic and
coloured by `fill`. Slices stay circular on any panel shape, so they look right
without `coord_fixed()`; pair it with `theme_void()` to drop the unused axes.

```{r pie}
shares <- data.frame(
  group  = c("Sketch", "Polish", "Coffee", "Doubt"),
  amount = c(40, 25, 20, 15)
)

ggplot(shares, aes(amount = amount, fill = group)) +
  geom_sketch_pie(seed = 1L) +
  scale_fill_sketch() +
  labs(title = "Where the time goes") +
  coord_fixed() +
  theme_void()
```

`geom_sketch_donut()` is the same with a hole; any `fill_style` hatches the
slices instead of filling them solid.

```{r donut}
ggplot(shares, aes(amount = amount, fill = group)) +
  geom_sketch_donut(fill_style = "hachure", seed = 2L) +
  scale_fill_sketch() +
  theme_void()
```

### Rounded bars

Rectangular geoms (`geom_sketch_rect()`, `geom_sketch_tile()`,
`geom_sketch_col()`/`bar()`) take a `corner_radius` for rounded corners — a
fraction of each half-side, so `0` is square and `1` is fully rounded.

```{r rounded-col}
ggplot(shares, aes(group, amount, fill = group)) +
  geom_sketch_col(corner_radius = 0.25, fill_style = "solid", seed = 1L,
                  show.legend = FALSE) +
  scale_fill_sketch() +
  labs(title = "Rounded columns", x = NULL) +
  theme_sketch()
```

### Content-aware arrows

`annotate_sketch_arrow()` points at a feature with a hand-drawn arrow. It is
*content-aware*: the shaft curves automatically toward the target, the
arrowhead orients to the curve's end tangent, and the label sits clear of the
shaft. A number for `curvature` (or `0` for straight) overrides the automatic
bow, and `arrow_type = "closed"` gives a filled head.

```{r arrow}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(colour = "grey30", seed = 1L) +
  annotate_sketch_arrow(x = 4.1, y = 33, xend = 5.25, yend = 18,
                        label = "heavy & thirsty", colour = "#C0392B",
                        seed = 2L) +
  annotate_sketch_arrow(x = 2.2, y = 12, xend = 1.7, yend = 30,
                        label = "light & frugal", colour = "#1F618D",
                        arrow_type = "closed", seed = 3L) +
  labs(title = "Pointing things out") +
  theme_sketch()
```

### Callouts

`annotate_sketch_callout()` puts a handwriting note in a roughened rounded box
that auto-sizes to the text, with a leader arrow to the target. Omit
`xend`/`yend` for a plain boxed label.

```{r callout}
ggplot(faithful, aes(eruptions, waiting)) +
  geom_sketch_point(colour = "grey40", seed = 1L) +
  annotate_sketch_callout(x = 2.1, y = 95, label = "short bursts",
                          xend = 1.9, yend = 75, fill = "#EAF2F8",
                          colour = "#1F618D", seed = 2L) +
  labs(title = "Boxed callouts") +
  theme_sketch()
```

### Hull marks

`geom_sketch_mark_hull()` circles a group of points with a roughened hull — the
sketch take on `ggforce::geom_mark_hull()`. Map `group`/`colour`/`fill` to mark
each cluster; with a `fill` the hull is shaded, otherwise it is outline-only.

```{r mark-hull}
ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_sketch_mark_hull(aes(fill = Species), expand = 0.08, seed = 1L) +
  geom_sketch_point(seed = 2L) +
  scale_colour_sketch() +
  scale_fill_sketch() +
  labs(title = "Grouping clusters") +
  theme_sketch()
```

### Bounding marks

`geom_sketch_mark_ellipse()`, `geom_sketch_mark_circle()`, and
`geom_sketch_mark_rect()` complete the family — a roughened ellipse, circle, or
rectangle around each group. The panel expands to fit the mark, so it is never
clipped.

```{r mark-ellipse}
ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_sketch_mark_ellipse(aes(fill = Species), seed = 1L) +
  geom_sketch_point(seed = 2L) +
  scale_colour_sketch() +
  scale_fill_sketch() +
  labs(title = "Bounding ellipses") +
  theme_sketch()
```

```{r mark-rect}
ggplot(iris, aes(Sepal.Length, Sepal.Width, colour = Species)) +
  geom_sketch_mark_rect(seed = 3L) +
  geom_sketch_point(seed = 2L) +
  scale_colour_sketch() +
  labs(title = "Bounding rectangles") +
  theme_sketch()
```

# New in 2.0

ggsketch 2.0 grows in two directions: new chart families (all built from the
same roughened grobs, so they inherit every fill style for free) and a
*drawing-medium simulator* — strokes that imitate pencil, ink, brush, charcoal,
marker and crayon, watercolour washes, textured papers, and hand-drawn coords.

## Dumbbell and slope charts

`geom_sketch_dumbbell()` draws a connector between two values per row, capped
with a sketch dot at each end — ideal for before/after comparisons.

```{r dumbbell}
dumb <- data.frame(g = c("Alpha", "Bravo", "Charlie", "Delta"),
                   before = c(20, 35, 28, 42),
                   after  = c(34, 51, 22, 47))
ggplot(dumb, aes(x = before, xend = after, y = reorder(g, after))) +
  geom_sketch_dumbbell(colour_x = "#B03A2E", colour_xend = "#1F618D",
                       seed = 1L) +
  labs(title = "Before vs after", x = "value", y = NULL) +
  theme_sketch()
```

`geom_sketch_slope()` connects each group across two (or more) categories, for
showing how a ranking changes.

```{r slope}
slope <- data.frame(
  time  = rep(c("Before", "After"), each = 4),
  value = c(20, 35, 28, 42, 34, 51, 22, 47),
  who   = rep(c("Alpha", "Bravo", "Charlie", "Delta"), 2)
)
ggplot(slope, aes(time, value, group = who, colour = who)) +
  geom_sketch_slope(seed = 1L) +
  scale_colour_sketch() +
  labs(title = "Slope chart", x = NULL) +
  theme_sketch()
```

## Beeswarm

`geom_sketch_beeswarm()` spreads points sideways so none overlap, keeping their
exact value — a deterministic, seeded swarm.

```{r beeswarm}
ggplot(iris, aes(Species, Sepal.Length, colour = Species)) +
  geom_sketch_beeswarm(size = 2.5, seed = 1L, show.legend = FALSE) +
  scale_colour_sketch() +
  labs(title = "geom_sketch_beeswarm()", x = NULL) +
  theme_sketch()
```

## Ridgelines

`geom_sketch_ridgeline()` stacks per-group densities into overlapping ridges,
drawn back-to-front. Any `fill_style` works, including watercolour.

```{r ridgeline}
ggplot(iris, aes(Sepal.Length, Species, fill = Species)) +
  geom_sketch_ridgeline(scale = 1.6, seed = 1L, show.legend = FALSE) +
  scale_fill_sketch() +
  labs(title = "geom_sketch_ridgeline()", y = NULL) +
  theme_sketch()
```

## Streamgraphs

`geom_sketch_streamgraph()` stacks group values around a flowing baseline
(`offset = "silhouette"`, `"zero"`, or `"wiggle"`).

```{r streamgraph}
set.seed(1)
stream <- expand.grid(t = 1:12, grp = c("a", "b", "c", "d"))
stream$v <- abs(sin(stream$t / 3 + match(stream$grp, letters)) + 1.2) * 5
ggplot(stream, aes(t, v, fill = grp)) +
  geom_sketch_streamgraph(seed = 1L) +
  scale_fill_sketch() +
  labs(title = "geom_sketch_streamgraph()", x = "time", y = NULL) +
  theme_sketch()
```

## Waffle and treemap

`geom_sketch_waffle()` turns counts into a grid of squares (largest-remainder
rounding to 100 cells by default).

```{r waffle}
waf <- data.frame(grp = c("Rent", "Food", "Travel", "Other"),
                  spend = c(45, 25, 20, 10))
ggplot(waf, aes(fill = grp, weight = spend)) +
  geom_sketch_waffle(seed = 1L) +
  scale_fill_sketch() +
  coord_equal() +
  labs(title = "geom_sketch_waffle()") +
  theme_void()
```

`geom_sketch_treemap()` lays out nested rectangles by `area` (a squarified
treemap) and can label each tile.

```{r treemap}
tm <- data.frame(grp = c("Alpha", "Bravo", "Charlie", "Delta", "Echo"),
                 val = c(40, 25, 15, 12, 8))
ggplot(tm, aes(area = val, fill = grp, label = grp)) +
  geom_sketch_treemap(seed = 1L, show.legend = FALSE) +
  scale_fill_sketch() +
  coord_equal() +
  labs(title = "geom_sketch_treemap()") +
  theme_void()
```

## Calendar heatmaps

`geom_sketch_calendar()` maps a `date` aesthetic onto a GitHub-style grid of
weeks (columns) and weekdays (rows).

```{r calendar}
set.seed(1)
cal <- data.frame(day = as.Date("2024-01-01") + 0:180)
cal$value <- cumsum(rnorm(nrow(cal)))
ggplot(cal, aes(date = day, fill = value)) +
  geom_sketch_calendar(seed = 1L) +
  scale_fill_viridis_c() +
  coord_equal() +
  labs(title = "geom_sketch_calendar()") +
  theme_sketch() +
  theme(axis.title = element_blank())
```

## Hand-drawn networks

`sketch_graph()` turns an edge list (or an **igraph** object) into ready-to-plot
`nodes` and `edges` frames, placing the nodes with a pure-R force-directed
layout — so the feature needs no graph dependency. Draw the result with
`geom_sketch_edge()` (roughened, optionally curved connectors) and
`geom_sketch_node()` (roughened markers with optional handwriting labels).

```{r network-basic, fig.height = 5}
edges <- data.frame(
  from = c("A", "A", "A", "B", "C", "C", "D", "E", "E", "F", "B", "G"),
  to   = c("B", "C", "D", "C", "D", "E", "E", "F", "G", "G", "E", "H")
)
g <- sketch_graph(edges, seed = 1L)

ggplot() +
  geom_sketch_edge(data = g$edges,
                   aes(x = x, y = y, xend = xend, yend = yend),
                   colour = "grey55", seed = 1L) +
  geom_sketch_node(data = g$nodes, aes(x = x, y = y, label = name),
                   size = 7, colour = "#1F618D", seed = 2L) +
  coord_equal() +
  labs(title = "A hand-drawn network") +
  theme_void()
```

Any node or edge column carried through `sketch_graph()` maps like a normal
aesthetic, and `curvature` bows the edges. Here node size encodes degree:

```{r network-curved, fig.height = 5}
deg <- table(c(edges$from, edges$to))
g2  <- sketch_graph(edges, seed = 4L)
g2$nodes$degree <- as.integer(deg[g2$nodes$name])

ggplot() +
  geom_sketch_edge(data = g2$edges,
                   aes(x = x, y = y, xend = xend, yend = yend),
                   curvature = 0.25, colour = "#B9770E", seed = 3L) +
  geom_sketch_node(data = g2$nodes,
                   aes(x = x, y = y, size = degree, label = name),
                   colour = "#7D3C98", show.legend = FALSE, seed = 4L) +
  scale_size_area(max_size = 13) +
  coord_equal() +
  labs(title = "Curved edges, nodes sized by degree") +
  theme_void()
```

Pass an **igraph** object straight to `sketch_graph()` for richer generators and
graph algorithms:

```{r network-igraph, eval = requireNamespace("igraph", quietly = TRUE), fig.height = 5}
set.seed(1)
ig <- igraph::sample_pa(16, directed = FALSE)
g3 <- sketch_graph(ig, seed = 7L)

ggplot() +
  geom_sketch_edge(data = g3$edges,
                   aes(x = x, y = y, xend = xend, yend = yend),
                   colour = "grey60", seed = 5L) +
  geom_sketch_node(data = g3$nodes, aes(x = x, y = y),
                   size = 5, colour = "#148F77", seed = 6L) +
  coord_equal() +
  labs(title = "A preferential-attachment graph (via igraph)") +
  theme_void()
```

## Hand-drawn maps (sf)

`geom_sketch_sf()` is a sketch take on `geom_sf()`: in one call it roughens
whichever simple-features geometry is present — `(MULTI)POLYGON` features get a
hole-aware hachure (or any `fill_style`), `(MULTI)LINESTRING` features become
sketch paths, and `(MULTI)POINT` features become sketch points. It needs the
optional **sf** package and plots in planar coordinates (pre-project lon/lat
data with `sf::st_transform()` for a faithful map).

```{r sf-choropleth, eval = requireNamespace("sf", quietly = TRUE), fig.height = 3.4}
nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)

ggplot() +
  geom_sketch_sf(data = nc, aes(fill = BIR74), seed = 1L) +
  scale_fill_viridis_c(option = "magma") +
  labs(title = "North Carolina births, 1974") +
  theme_void()
```

Switch the `fill_style`, exactly like any other sketch fill:

```{r sf-crosshatch, eval = requireNamespace("sf", quietly = TRUE), fig.height = 3.4}
ggplot() +
  geom_sketch_sf(data = nc, aes(fill = SID74), fill_style = "cross_hatch",
                 seed = 2L) +
  scale_fill_distiller(palette = "RdPu", direction = 1) +
  labs(title = "Cross-hatched choropleth of SIDS cases") +
  theme_void()
```

## Radar charts

`geom_sketch_radar()` draws a spider chart: each series is a closed polygon over
evenly spaced axes, with a roughened web (rings, spokes, and labels) behind. Map
`axis`, `value`, `group`, and `colour`/`fill`; like the pie geoms it lives in
its own square space, so pair it with `coord_equal()` and `theme_void()`.

```{r radar, fig.height = 5}
skills <- data.frame(
  axis  = rep(c("Speed", "Power", "Range", "Control", "Stamina", "Magic"), 2),
  value = c(8, 6, 9, 5, 7, 4, 5, 9, 4, 8, 6, 9),
  who   = rep(c("Aria", "Bilo"), each = 6)
)
ggplot(skills, aes(axis = axis, value = value, group = who,
                   colour = who, fill = who)) +
  geom_sketch_radar(alpha = 0.3, seed = 1L) +
  scale_colour_sketch() +
  scale_fill_sketch() +
  coord_equal() +
  labs(title = "Character stats") +
  theme_void()
```

A single series reads cleanly with a watercolour wash and more grid rings:

```{r radar-one, fig.height = 5}
ggplot(subset(skills, who == "Aria"),
       aes(axis = axis, value = value, group = who)) +
  geom_sketch_radar(fill = "#2E86C1", fill_style = "watercolor",
                    n_rings = 5, seed = 3L) +
  coord_equal() +
  labs(title = "One series, watercolour") +
  theme_void()
```

## Chord diagrams

`geom_sketch_chord()` lays nodes on a circle — each given a rim arc sized by its
total flow — and draws every weighted relation as a ribbon through the centre.
Give it an edge table and the `from`, `to`, and `value` columns; ribbons colour
by source node.

```{r chord, fig.height = 5.4}
trade <- data.frame(
  from  = c("Asia", "Asia", "Europe", "Africa", "Africa", "America",
            "America", "Asia"),
  to    = c("Europe", "America", "America", "Asia", "Europe", "Africa",
            "Europe", "Africa"),
  value = c(8, 6, 5, 3, 2, 4, 7, 5)
)
ggplot() +
  geom_sketch_chord(trade, from, to, value, seed = 1L) +
  scale_fill_sketch() +
  coord_equal() +
  labs(title = "Trade flows") +
  theme_void()
```

## Arc diagrams

`geom_sketch_arc_diagram()` is a linear cousin of the chord diagram: nodes sit on
a horizontal line and each weighted relation arcs over the axis as a roughened
semicircle, coloured by source and thickened by `value`.

```{r arc-diagram, fig.height = 3.8}
rel <- data.frame(
  from  = c("Anna", "Anna", "Ben", "Cara", "Cara", "Dan", "Eve", "Anna"),
  to    = c("Ben", "Cara", "Cara", "Dan", "Eve", "Eve", "Anna", "Dan"),
  value = c(3, 1, 2, 4, 2, 1, 3, 2)
)
ggplot() +
  geom_sketch_arc_diagram(rel, from, to, value, seed = 5L, max_linewidth = 3) +
  scale_colour_sketch() +
  labs(title = "Who talks to whom") +
  theme_void() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
```

## Dendrograms

`geom_sketch_dendrogram()` draws a hierarchical-clustering tree — the right-angle
elbows roughened into a hand-drawn wobble. Pass an `hclust` object or a numeric
data frame (it clusters for you).

```{r dendrogram, fig.height = 4.6}
cars12 <- mtcars[1:12, c("mpg", "disp", "hp", "wt", "qsec")]
ggplot() +
  geom_sketch_dendrogram(cars12, seed = 3L, line_width = 1) +
  labs(title = "Clustering cars") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(8, 8, 28, 14))
```

## Bump (ranking) charts

`geom_sketch_bump()` draws each series' rank at every time point, joined across
adjacent times by smooth roughened curves so a crossing reads as an overtake.
Give it long data with the time, series, and value columns.

```{r bump, fig.height = 4.4}
standings <- data.frame(
  year = rep(2018:2023, each = 5),
  team = rep(c("Falcons", "Bears", "Wolves", "Hawks", "Lions"), times = 6),
  pts  = c(20, 18, 15, 12, 10,  15, 20, 18, 10, 12,  18, 12, 20, 15, 8,
           12, 18, 15, 20, 14,  10, 15, 12, 18, 22,  22, 14, 10, 16, 18)
)
ggplot() +
  geom_sketch_bump(standings, year, team, pts, seed = 4L, point_size = 4) +
  scale_colour_sketch() +
  labs(title = "Season standings") +
  theme_void() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
```

## Alluvial / Sankey diagrams

`geom_sketch_alluvial()` draws two or more categorical axes as stacks of strata,
joined by flows whose thickness is the frequency of each category combination.
Give it a wide data frame, the `axes` columns in order, and an optional `value`
weight; flows colour by the first axis (or a named `fill` column).

```{r alluvial, fig.height = 5}
titanic <- as.data.frame(Titanic)
ggplot() +
  geom_sketch_alluvial(titanic, axes = c("Class", "Sex", "Age", "Survived"),
                       value = "Freq", seed = 1L) +
  scale_fill_sketch() +
  labs(title = "Titanic passengers") +
  theme_void()
```

## Parallel coordinates

`geom_sketch_parallel()` draws several numeric columns as vertical axes and
every observation as a roughened polyline crossing them. Axes scale
independently; map `colour` to a grouping column.

```{r parallel, fig.height = 4.6}
ggplot() +
  geom_sketch_parallel(iris,
    axes = c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width"),
    colour = "Species", alpha = 0.6, seed = 1L) +
  scale_colour_sketch() +
  labs(title = "Iris parallel coordinates") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Mosaic plots

`geom_sketch_mosaic()` splits the square into columns by the marginal counts of
`x`, then each column by the conditional counts of `y`, so every tile's area is
the joint frequency.

```{r mosaic, fig.height = 4.8}
titanic <- as.data.frame(Titanic)
ggplot() +
  geom_sketch_mosaic(titanic, x = Class, y = Survived, value = Freq,
                     seed = 1L) +
  scale_fill_sketch() +
  labs(title = "Survival by class") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Coxcomb / Nightingale rose

`geom_sketch_rose()` gives each category an equal angular wedge whose radius (or,
with `area_true = TRUE`, whose *area*) encodes the value — Florence Nightingale's
coxcomb. An optional `fill` stacks radially within each wedge.

```{r rose, fig.height = 5.4}
mortality <- data.frame(
  quarter = rep(c("Q1", "Q2", "Q3", "Q4", "Q5", "Q6"), each = 3),
  cause   = rep(c("Disease", "Wounds", "Other"), times = 6),
  n       = c(20, 8, 4,  30, 10, 5,  18, 6, 3,  12, 9, 2,  8, 5, 2,  14, 7, 3)
)
ggplot() +
  geom_sketch_rose(mortality, quarter, n, fill = cause, area_true = TRUE,
                   seed = 8L, alpha = 0.85) +
  scale_fill_sketch() +
  labs(title = "Mortality by quarter (area-true)", fill = NULL) +
  coord_equal() +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Marimekko charts

`geom_sketch_marimekko()` draws variable-width stacked bars: column width is one
category's share of the total, the stacked segments are a second category's
shares, and each tile's area is the joint value. Width percentages sit on top.

```{r marimekko, fig.height = 4.6}
sales <- data.frame(
  region  = rep(c("North America", "Europe", "Asia", "Other"), each = 3),
  product = rep(c("Phones", "Laptops", "Tablets"), times = 4),
  revenue = c(50, 35, 15,  30, 40, 20,  45, 25, 30,  10, 12, 6)
)
ggplot() +
  geom_sketch_marimekko(sales, region, product, revenue, seed = 6L) +
  scale_fill_sketch() +
  labs(title = "Revenue share by region and product", fill = NULL) +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.margin = margin(10, 10, 18, 10))
```

## Sunburst charts

`geom_sketch_sunburst()` draws a hierarchy as nested rings of annular sectors:
the columns in `levels` go from the inner root ring outward, and each deeper ring
splits its parent's angle by the children's summed `value`, so a child always
nests inside its parent. Fill by the top-level ancestor for the classic look.

```{r sunburst, fig.height = 5.6}
gss <- data.frame(
  region = rep(c("West", "East", "North"), each = 4),
  dept   = rep(c("Sales", "Sales", "Eng", "Eng"), 3),
  team   = paste0("T", 1:12),
  n      = c(6, 3, 8, 2, 5, 4, 3, 7, 2, 6, 4, 5)
)
ggplot() +
  geom_sketch_sunburst(gss, levels = c("region", "dept", "team"), value = "n",
                       fill_by = "root", label = TRUE, label_size = 2.6,
                       seed = 7L) +
  scale_fill_sketch() +
  coord_equal() +
  labs(title = "Headcount by region › dept › team") +
  theme_void() +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))
```

## Animation: boiling lines

`animate_sketch()` animates a sketch plot two ways. `type = "boil"` re-renders
it while shifting every roughening seed per frame, so the whole drawing shimmers
like a hand-animated cel; `type = "draw_on"` reveals the finished drawing behind
a moving wipe, as if a hand were drawing it on. Frames are stitched into a GIF
when **gifski** or **magick** is installed; otherwise the frame paths are
returned. It animates any sketch plot with no change to its code.

```{r animate, eval = requireNamespace("magick", quietly = TRUE) || requireNamespace("gifski", quietly = TRUE)}
p <- ggplot(mpg, aes(class)) +
  geom_sketch_bar(fill = "#7BAFD4", seed = 1L) +
  labs(title = "Boiling bars", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))

gif <- animate_sketch(p, nframes = 10, fps = 8,
                      file = knitr::fig_path(".gif"),
                      width = 6, height = 4, res = 110)
knitr::include_graphics(gif)
```

And the same plot drawn on, left to right:

```{r animate-drawon, eval = requireNamespace("magick", quietly = TRUE) || requireNamespace("gifski", quietly = TRUE)}
drawn <- animate_sketch(p, type = "draw_on", nframes = 14, fps = 10,
                        file = knitr::fig_path(".gif"),
                        width = 6, height = 4, res = 110)
knitr::include_graphics(drawn)
```

## Drawing media

The `medium` controls *how* a stroke is laid down. `sketch_media()` lists the
seven media; set one as a constant on a line/path/segment geom, or **map** it as
an aesthetic with `scale_medium_discrete()`.

```{r media-list}
sketch_media()
```

```{r medium-aes}
lv <- c("pencil", "ink", "brush", "charcoal", "marker")
waves <- data.frame(x = rep(1:40, length(lv)),
                    g = factor(rep(lv, each = 40), levels = lv))
waves$y <- as.integer(waves$g) + sin(waves$x / 4) * 0.3
ggplot(waves, aes(x, y, group = g, medium = g, colour = g)) +
  geom_sketch_line(linewidth = 1.1, seed = 1L) +
  scale_medium_discrete(media = lv) +
  scale_colour_sketch() +
  labs(title = "medium mapped to a variable", x = NULL, y = NULL) +
  theme_sketch()
```

## Watercolour fills

`fill_style = "watercolor"` paints stacked translucent washes with soft bleeds
and granulation, on any polygon-, ribbon-, ellipse-, or band-filled geom.

```{r watercolor-poly}
ang  <- seq(0, 2 * pi, length.out = 11)[-11]
r    <- rep(c(1, 0.45), length.out = 10)
star <- data.frame(x = r * cos(ang), y = r * sin(ang))
ggplot(star, aes(x, y)) +
  geom_sketch_polygon(fill = "#2E86C1", fill_style = "watercolor", seed = 1L) +
  coord_equal() +
  labs(title = "fill_style = \"watercolor\"") +
  theme_sketch()
```

```{r watercolor-violin}
ggplot(mpg, aes(class, hwy, fill = class)) +
  geom_sketch_violin(fill_style = "watercolor", seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Watercolour violins", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Textured paper

`theme_sketch(paper = )` draws the panel on a textured ground —
`sketch_papers()` lists them (notebook, graph, dotted, aged, blueprint,
chalkboard, kraft). Dark grounds flip the text light automatically.

```{r paper-list}
sketch_papers()
```

```{r paper-notebook}
ggplot(sales, aes(product, units)) +
  geom_sketch_col(fill = "#7BAFD4", seed = 1L) +
  labs(title = "Notebook paper", x = NULL) +
  theme_sketch(paper = "notebook")
```

```{r paper-blueprint}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(colour = "white", seed = 1L) +
  labs(title = "Blueprint ground") +
  theme_sketch(paper = "blueprint")
```

You can also drop a paper onto any theme via `element_sketch_paper()`:

```{r paper-element}
ggplot(faithful, aes(eruptions, waiting)) +
  geom_sketch_point(colour = "#1F618D", seed = 1L) +
  theme_sketch() +
  theme(panel.background = element_sketch_paper("graph"))
```

## Hand-drawn coordinate frames

`coord_sketch()` roughens the gridlines and ticks under *any* theme — no need
for a sketch theme at all:

```{r coord-sketch}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(seed = 1L) +
  labs(title = "coord_sketch() under a plain theme") +
  coord_sketch(seed = 1L)
```

`coord_sketch_polar()` is the polar companion — a wobbly circular grid for rose
and pie-style charts:

```{r coord-polar}
rose <- data.frame(g = c("a", "b", "c", "d", "e", "f"),
                   v = c(3, 5, 2, 4, 6, 3))
ggplot(rose, aes(g, v, fill = g)) +
  geom_sketch_col(seed = 1L, show.legend = FALSE) +
  scale_fill_sketch() +
  coord_sketch_polar(seed = 1L) +
  labs(title = "coord_sketch_polar()", x = NULL, y = NULL) +
  theme_sketch()
```

## Composition: facets, scales, coords

Sketch geoms respect the full grammar.

```{r facet}
ggplot(mpg, aes(displ, hwy)) +
  geom_sketch_point(size = 2.5, colour = "#34495E", seed = 9L) +
  geom_sketch_smooth(method = "lm", formula = y ~ x, seed = 10L) +
  facet_wrap(~drv, labeller = label_both) +
  labs(title = "Faceted by drivetrain") +
  theme_sketch()
```

## Dark mode

Every example above works with `theme_sketch(dark = TRUE)`:

```{r dark}
ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Dark preset", x = NULL) +
  theme_sketch(dark = TRUE)
```

## A hand-drawn frame

By default `theme_sketch()` keeps the gridlines, panel border, and axis ticks
crisp. Pass `rough_frame = TRUE` and the frame is roughened too, so it matches
the marks.

```{r rough-frame}
ggplot(sales, aes(product, units)) +
  geom_sketch_col(fill = "#7BAFD4", seed = 1L) +
  labs(title = "Everything wobbles", x = NULL) +
  theme_sketch(rough_frame = TRUE, seed = 1L)
```

The roughened elements are real theme elements — `element_sketch_line()` and
`element_sketch_rect()` — so you can also drop them into any theme yourself and
tune their `roughness`, `bowing`, and `seed`:

```{r element-sketch}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(seed = 1L) +
  theme_sketch() +
  theme(
    panel.grid.major = element_sketch_line(roughness = 0.8, seed = 7L),
    axis.ticks       = element_sketch_line(roughness = 0.6, seed = 8L)
  )
```

## A matching palette

`scale_colour_sketch()` / `scale_fill_sketch()` use a qualitative palette
(`sketch_palette()`) chosen to suit the hand-drawn look:

```{r scale-discrete}
ggplot(mpg, aes(displ, hwy, colour = drv)) +
  geom_sketch_point(size = 2.5, seed = 1L) +
  scale_colour_sketch() +
  labs(title = "scale_colour_sketch()") +
  theme_sketch(rough_frame = TRUE, seed = 2L)
```

For continuous data the `*_sketch_c()` variants give an ink-on-paper gradient:

```{r scale-continuous}
ggplot(faithful, aes(eruptions, waiting, colour = waiting)) +
  geom_sketch_point(size = 2.5, seed = 1L) +
  scale_colour_sketch_c() +
  labs(title = "scale_colour_sketch_c()") +
  theme_sketch()
```

## The scribble fill

`"scribble"` is one continuous winding stroke that overshoots the boundary, like
scribbling to fill a shape:

```{r scribble}
ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(fill_style = "scribble", seed = 3L, show.legend = FALSE) +
  scale_fill_sketch() +
  labs(title = "fill_style = \"scribble\"", x = NULL) +
  theme_sketch()
```

It works anywhere a `fill_style` is accepted. The eight styles:

```{r fill-styles}
styles <- c("hachure", "cross_hatch", "zigzag", "zigzag_line",
            "scribble", "dots", "dashed", "solid")
grid <- expand.grid(col = 1:4, row = 1:2)
grid$style <- styles
ggplot(grid) +
  lapply(seq_len(nrow(grid)), function(i) {
    geom_sketch_rect(
      data = grid[i, ],
      aes(xmin = col - 0.45, xmax = col + 0.45,
          ymin = row - 0.4,  ymax = row + 0.4),
      fill = "#7BAFD4", fill_style = grid$style[i], seed = i
    )
  }) +
  geom_sketch_text(aes(col, row - 0.55, label = style), size = 4) +
  coord_equal() +
  labs(title = "The eight fill styles", x = NULL, y = NULL) +
  theme_sketch() +
  theme(axis.text = element_blank())
```

## Reproducible handwriting fonts

`geom_sketch_text()` picks up a handwriting face preinstalled on your OS, but for
results that reproduce on any machine or CI runner, register a font file
explicitly with `register_sketch_font()` and a font-aware device (ragg, svglite,
cairo):

```r
register_sketch_font("Caveat", "path/to/Caveat-Regular.ttf")

ggplot(lab, aes(x, y, label = txt)) +
  geom_sketch_text(family = "Caveat", size = 8) +
  theme_sketch()
```
