---
title: "Gallery — every geom"
description: >
  A tour of every ggsketch geom with runnable examples, from bars and lines to
  boxplots, ribbons, circles, and annotations.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  message    = FALSE,
  warning    = FALSE,
  fig.width  = 7,
  fig.height = 4.2,
  dpi        = 120,
  dev        = "ragg_png",
  out.width  = "100%",
  fig.align  = "center"
)
library(ggplot2)
library(ggsketch)
set.seed(1)
options(ggsketch.seed = 1)
```

Everything here is a real ggplot2 layer, so it composes with `aes()`, stats,
scales, facets, and coords — and renders on any device. Every example sets a
`seed` so the wobble is reproducible.

## Bars and columns

`geom_sketch_col()` is the signature geom: a roughened outline with a hachure
(pencil-shading) fill.

```{r col-basic}
sales <- data.frame(product = c("Alpha", "Bravo", "Charlie", "Delta", "Echo"),
                    units   = c(34, 51, 22, 47, 39))

ggplot(sales, aes(product, units)) +
  geom_sketch_col(fill = "#7BAFD4", seed = 1L) +
  labs(title = "Units sold", x = NULL) +
  theme_sketch()
```

Map `fill` to a variable like any ggplot2 bar. Each bar gets its own seed offset,
so no two bars wobble identically.

```{r col-mapped}
ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(seed = 2L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Mapped fill", x = NULL) +
  theme_sketch()
```

`geom_sketch_bar()` counts rows for you (like `geom_bar()`):

```{r bar-count}
ggplot(mpg, aes(class)) +
  geom_sketch_bar(fill = "#C39BD3", seed = 3L) +
  labs(title = "Vehicle count by class", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

Bars flip and stack just like the originals:

```{r col-flip}
ggplot(sales, aes(reorder(product, units), units)) +
  geom_sketch_col(fill = "#F1948A", seed = 1L) +
  coord_flip() +
  labs(title = "Horizontal bars", x = NULL) +
  theme_sketch()
```

## Lines, paths, and points

```{r line}
econ <- economics[economics$date > as.Date("2000-01-01"), ]
ggplot(econ, aes(date, unemploy)) +
  geom_sketch_line(colour = "steelblue", linewidth = 0.8, seed = 1L) +
  labs(title = "US unemployment", x = NULL, y = "thousands") +
  theme_sketch()
```

`geom_sketch_point()` draws each point as a small roughened ellipse:

```{r point}
ggplot(mtcars, aes(wt, mpg, colour = factor(cyl))) +
  geom_sketch_point(size = 4, seed = 1L) +
  scale_colour_brewer("cylinders", palette = "Dark2") +
  labs(title = "Fuel economy vs weight") +
  theme_sketch()
```

Lines and points compose like any layers:

```{r line-point}
df <- data.frame(x = 1:12, y = c(3, 5, 4, 7, 6, 9, 8, 11, 9, 12, 11, 14))
ggplot(df, aes(x, y)) +
  geom_sketch_line(colour = "grey40", seed = 3L) +
  geom_sketch_point(size = 4, colour = "firebrick", seed = 8L) +
  labs(title = "Trend with markers") +
  theme_sketch()
```

Multiple groups, one seed per group:

```{r line-groups}
ggplot(ggplot2::economics_long, aes(date, value01, colour = variable)) +
  geom_sketch_line(seed = 5L) +
  labs(title = "Five series, hand-drawn", x = NULL, y = NULL) +
  theme_sketch()
```

## Rectangles and tiles

```{r rect}
rects <- data.frame(xmin = c(1, 3, 5), xmax = c(2, 4, 6),
                    ymin = 0, ymax = c(2, 4, 3))
ggplot(rects) +
  geom_sketch_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
                       fill = factor(xmin)),
                   seed = 1L, show.legend = FALSE) +
  labs(title = "geom_sketch_rect()") +
  theme_sketch()
```

A sketchy heatmap with `geom_sketch_tile()`:

```{r tile}
td <- expand.grid(x = 1:8, y = 1:6)
td$z <- td$x + td$y
ggplot(td, aes(x, y, fill = z)) +
  geom_sketch_tile(seed = 2L, hachure_gap = 0.18) +
  scale_fill_viridis_c() +
  labs(title = "geom_sketch_tile()") +
  theme_sketch()
```

## Polygons, ribbons, areas, and densities

Concave polygons fill correctly (the hachure respects every notch):

```{r polygon}
ang  <- seq(0, 2 * pi, length.out = 11)[-11]
r    <- rep(c(1, 0.45), length.out = 10)
star <- data.frame(x = r * cos(ang), y = r * sin(ang))
ggplot(star, aes(x, y)) +
  geom_sketch_polygon(fill = "tomato", seed = 1L) +
  coord_equal() +
  labs(title = "A concave star") +
  theme_sketch()
```

```{r ribbon-area}
band <- data.frame(x = 1:20)
band$y  <- 10 + 5 * sin(seq(0, 3 * pi, length.out = 20))
band$lo <- band$y - 2
band$hi <- band$y + 2

ggplot(band, aes(x)) +
  geom_sketch_ribbon(aes(ymin = lo, ymax = hi), fill = "plum", seed = 2L) +
  geom_sketch_line(aes(y = y), seed = 3L) +
  labs(title = "Ribbon + line") +
  theme_sketch()
```

```{r area}
ggplot(band, aes(x, y)) +
  geom_sketch_area(fill = "lightgreen", seed = 3L) +
  labs(title = "geom_sketch_area()") +
  theme_sketch()
```

```{r density}
ggplot(faithful, aes(eruptions)) +
  geom_sketch_density(fill = "khaki", seed = 4L) +
  labs(title = "Old Faithful eruptions") +
  theme_sketch()
```

## Smooths

A hand-drawn fit with a roughened confidence band:

```{r smooth}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(seed = 1L) +
  geom_sketch_smooth(method = "lm", formula = y ~ x, seed = 2L) +
  labs(title = "Linear fit with CI band") +
  theme_sketch()
```

```{r smooth-loess}
ggplot(mpg, aes(displ, hwy)) +
  geom_sketch_point(colour = "grey50", seed = 1L) +
  geom_sketch_smooth(seed = 2L, colour = "darkorange") +
  labs(title = "loess fit") +
  theme_sketch()
```

## Circles and ellipses

Radii are in data units, so use `coord_equal()` for true circles:

```{r circle}
cdf <- data.frame(x = c(1, 3, 2), y = c(1, 1, 2.5),
                  r = c(0.6, 0.9, 0.5), grp = c("a", "b", "c"))
ggplot(cdf, aes(x, y, r = r, fill = grp)) +
  geom_sketch_circle(seed = 1L, show.legend = FALSE) +
  coord_equal() +
  labs(title = "geom_sketch_circle()") +
  theme_sketch()
```

```{r ellipse}
edf <- data.frame(x = c(1, 3), y = c(1, 2), a = c(1.4, 0.8), b = c(0.6, 1.2))
ggplot(edf, aes(x, y, a = a, b = b, fill = factor(x))) +
  geom_sketch_ellipse(seed = 2L, show.legend = FALSE) +
  coord_equal() +
  labs(title = "geom_sketch_ellipse()") +
  theme_sketch()
```

## Segments and steps

```{r segment}
sdf <- data.frame(x = 1:4, y = c(1, 3, 2, 4),
                  xend = 2:5, yend = c(3, 1, 4, 2))
ggplot(sdf) +
  geom_sketch_segment(aes(x = x, y = y, xend = xend, yend = yend),
                      colour = "darkgreen", linewidth = 1, seed = 3L) +
  labs(title = "geom_sketch_segment()") +
  theme_sketch()
```

```{r step}
stp <- data.frame(x = 1:8, y = c(1, 3, 2, 5, 4, 6, 5, 8))
ggplot(stp, aes(x, y)) +
  geom_sketch_step(colour = "purple", linewidth = 1, seed = 4L) +
  geom_sketch_point(seed = 5L) +
  labs(title = "geom_sketch_step()") +
  theme_sketch()
```

## Boxplots

A composed geom: rough IQR box, thick median, whiskers, and sketchy outliers.

```{r boxplot}
ggplot(mpg, aes(class, hwy)) +
  geom_sketch_boxplot(seed = 1L) +
  labs(title = "Highway mpg by class", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

By default the box is `fill_style = "solid"` (outline only). For coloured,
shaded boxes, map `fill` and switch on a fill style:

```{r boxplot-hachure}
ggplot(mpg, aes(class, hwy, fill = class)) +
  geom_sketch_boxplot(fill_style = "hachure", seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(title = "Hachure-filled boxes", x = NULL) +
  theme_sketch() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Annotations

`annotate_sketch()` adds one-off hand-drawn marks (no `aes()` inheritance):

```{r annotate}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(seed = 1L) +
  annotate_sketch("rect", xmin = 3, xmax = 4, ymin = 15, ymax = 22,
                  fill = NA, colour = "red", seed = 2L) +
  annotate_sketch("segment", x = 2, y = 32, xend = 3.4, yend = 21,
                  colour = "blue", linewidth = 1, seed = 3L) +
  annotate_sketch("circle", x = 5, y = 30, r = 0.4,
                  colour = "darkgreen", fill = "green", seed = 4L) +
  labs(title = "Highlighting with annotate_sketch()") +
  theme_sketch()
```

## Composition: facets, scales, coords

Sketch geoms respect the full grammar.

```{r facet}
ggplot(mpg, aes(displ, hwy)) +
  geom_sketch_point(size = 2.5, colour = "#34495E", seed = 9L) +
  geom_sketch_smooth(method = "lm", formula = y ~ x, seed = 10L) +
  facet_wrap(~drv, labeller = label_both) +
  labs(title = "Faceted by drivetrain") +
  theme_sketch()
```

## Dark mode

Every example above works with `theme_sketch(dark = TRUE)`:

```{r dark}
ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Dark preset", x = NULL) +
  theme_sketch(dark = TRUE)
```
