---
title: "How it works"
description: >
  The three-layer architecture, why roughening happens in inch space, how
  determinism works, and the algorithm credits.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 4, dpi = 120, dev = "ragg_png",
  out.width = "100%", fig.align = "center"
)
library(ggplot2)
library(ggsketch)
options(ggsketch.seed = 1)
```

ggsketch is built in three strictly separated layers. Keeping them apart is what
makes the package testable, device-independent, and reproducible.

## Layer 1 — pure geometry

The `core-*` functions map **numbers to numbers**. They take coordinates and
sketch parameters and return roughened coordinates — no grid, no ggplot2, no
drawing. This is where the rough.js *algorithms* are reimplemented in original R:
point roughening, cubic-Bézier sampling, ellipse generation, and an
Active-Edge-Table scan-line filler for hachure (including concave polygons).

You can call them directly:

```{r layer1-line}
# Two passes of a roughened polyline through three points
passes <- roughen_polyline(x = c(0, 1, 2), y = c(0, 1, 0),
                           roughness = 1, n_passes = 2, seed = 1L)
length(passes)       # one matrix per pass
head(passes[[1]])    # x/y coordinates of the first stroke
```

```{r layer1-fill}
# Hachure fill of a unit square: a list of line segments
segs <- hachure_fill(c(0, 1, 1, 0), c(0, 0, 1, 1),
                     hachure_gap = 0.2, hachure_angle = 45, seed = 1L)
length(segs)
```

Because Layer 1 is pure, the package's primary regression tests are deterministic
**geometry snapshots** — they compare these numbers exactly, which is far more
stable than comparing rendered images.

## Layer 2 — grid grobs

The `*_grob()` constructors wrap Layer 1 in grid grobs whose `makeContent()`
method does the drawing. Crucially, roughening happens **after** converting to
absolute device inches:

```{r layer2}
g <- sketch_path_grob(x = c(0.1, 0.5, 0.9), y = c(0.2, 0.8, 0.5),
                      roughness = 1, seed = 1L)
class(g)
```

Working in inch space (not data or npc space) means the wobble has a consistent
physical scale regardless of the panel's aspect ratio or the output device — and
because `makeContent()` re-runs on each draw, resizing a plot re-roughens cleanly
at the new size instead of stretching a cached path.

## Layer 3 — ggproto geoms

The `geom_sketch_*()` functions are ordinary ggplot2 layers built with the public
extension API (`ggproto`, `draw_panel`/`draw_group`, `coord$transform`,
`draw_key`). That is why they compose with everything in the grammar:

```{r layer3}
ggplot(mtcars, aes(wt, mpg, colour = factor(cyl))) +
  geom_sketch_point(seed = 1L) +
  geom_sketch_smooth(method = "lm", formula = y ~ x, se = FALSE, seed = 2L) +
  facet_wrap(~am) +
  theme_sketch()
```

## Determinism

Every randomized routine draws from a **seeded local RNG stream** and restores
the global random state afterwards. So sketch plots are reproducible given a
`seed`, and they never disturb `set.seed()` elsewhere in your code:

```{r determinism}
set.seed(99)
before <- runif(1)

# draw something sketchy (consumes randomness internally)
invisible(roughen_polyline(c(0, 1), c(0, 1), seed = 1L))

set.seed(99)
after <- runif(1)
identical(before, after)   # TRUE — global RNG untouched
```

## Credits & non-affiliation

The algorithms are reimplemented in original R from the **published
descriptions** of the [rough.js](https://github.com/rough-stuff/rough)
algorithms ([Preet Shihn, 2020](https://shihn.ca/posts/2020/roughjs-algorithms/))
and the hachure approach of Wood et al. No rough.js source is vendored, copied,
or translated, and no JavaScript ships in the package.

> ggsketch is an independent R package reimplementing the hand-drawn sketch
> aesthetic from first principles. It is not affiliated with, derived from, or
> endorsed by the rough.js project, ggrough, or any related JavaScript libraries.

rough.js is © Preet Shihn and licensed MIT. ggsketch is licensed MIT. See
`inst/NOTICE` for the full attribution.
