---
title: "Theming & fonts"
description: >
  theme_sketch() light and dark presets, the muted colour palette, and optional
  handwriting fonts.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 4.2, dpi = 120, dev = "ragg_png",
  out.width = "100%", fig.align = "center"
)
library(ggplot2)
library(ggsketch)
options(ggsketch.seed = 1)
```

The sketchiness of the *marks* comes from the geoms. `theme_sketch()` styles the
surrounding frame — typography, gridlines, background — with a muted palette to
match.

## Light and dark presets

```{r light}
sales <- data.frame(product = c("Alpha", "Bravo", "Charlie", "Delta"),
                    units   = c(34, 51, 22, 47))

ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Light (paper) preset", x = NULL) +
  theme_sketch()
```

```{r dark}
ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(colour = "grey85", seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Dark (chalkboard) preset", x = NULL) +
  theme_sketch(dark = TRUE)
```

On a dark background, give the geoms a light outline `colour` (e.g. `"grey85"`)
so the rough strokes read clearly.

## Base size and the rest of the grammar

`theme_sketch()` is a normal ggplot2 theme — combine and override freely.

```{r size}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(size = 3, seed = 1L) +
  labs(title = "Bigger base text", subtitle = "theme_sketch(base_size = 15)") +
  theme_sketch(base_size = 15) +
  theme(panel.grid.minor = element_blank())
```

## Handwriting fonts (optional)

The look does **not** depend on fonts, but a handwriting face for the text adds
to it. Pass a family name to `base_family`, or use `"auto"` to pick the first
installed handwriting font (falling back to the device default if none are
found).

### What they look like

A few handwriting faces that pair well with the geoms — Google Fonts like
[Caveat](https://fonts.google.com/specimen/Caveat),
[Permanent Marker](https://fonts.google.com/specimen/Permanent+Marker), and
[Indie Flower](https://fonts.google.com/specimen/Indie+Flower), plus the
handwriting fonts that ship with Windows and macOS (Segoe Print, Ink Free,
Bradley Hand, Chalkboard, Comic Sans MS). The specimen below renders whichever
of those are installed on the build machine.

```{r specimen-build, include = FALSE}
have_font <- function(f) {
  if (!requireNamespace("systemfonts", quietly = TRUE)) return(FALSE)
  reg <- tryCatch(systemfonts::registry_fonts()$family, error = function(e) character())
  isTRUE(f %in% c(systemfonts::system_fonts()$family, reg))
}
candidates <- c("Caveat", "Permanent Marker", "Indie Flower", "Shadows Into Light",
                "Architects Daughter", "Segoe Print", "Ink Free", "Bradley Hand",
                "Chalkboard", "Comic Sans MS")
fonts <- Filter(have_font, candidates)
```

```{r specimen, echo = FALSE, eval = length(fonts) > 0, fig.height = 0.55 * length(fonts) + 0.8}
spec <- data.frame(y = rev(seq_along(fonts)), family = fonts)
ggplot(spec, aes(0.02, y)) +
  lapply(seq_len(nrow(spec)), function(i) {
    geom_sketch_text(
      data = spec[i, ],
      aes(label = paste0(family, " — sketch it by hand")),
      family = spec$family[i], hjust = 0, size = 6
    )
  }) +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0.5, nrow(spec) + 0.5)) +
  coord_cartesian(clip = "off") +
  labs(title = "Handwriting fonts found on this machine", x = NULL, y = NULL) +
  theme_sketch() +
  theme(axis.text = element_blank(), panel.grid = element_blank())
```

```{r specimen-none, echo = FALSE, results = "asis", eval = length(fonts) == 0}
cat("*No handwriting fonts were found on the build machine, so the specimen is",
    "skipped. Install one of the faces above (or register one — see below) to",
    "see it here.*")
```

### Using one

Pass `base_family = "auto"` to style the whole theme with the first installed
handwriting font; `geom_sketch_text()` picks it up the same way.

```{r auto, eval = length(fonts) > 0}
ggplot(sales, aes(product, units)) +
  geom_sketch_col(fill = "#7BAFD4", seed = 1L) +
  geom_sketch_text(aes(label = units), nudge_y = 2.5, size = 6) +
  labs(title = "base_family = \"auto\"", x = NULL) +
  theme_sketch(base_family = "auto")
```

Check what is available on your machine:

```{r check-fonts, eval = FALSE}
ggsketch_check_fonts()
#> Available handwriting fonts:
#>   Caveat
```

### Reproducible fonts

To get the *same* face on any machine or CI runner — without relying on a system
install — download a font once (e.g. Caveat from Google Fonts) and register it
with `register_sketch_font()`, then render with a font-aware device (ragg,
svglite, cairo):

```{r register, eval = FALSE}
register_sketch_font("Caveat", "~/fonts/Caveat-Regular.ttf")

ggplot(sales, aes(product, units)) +
  geom_sketch_col(fill = "#7BAFD4", seed = 1L) +
  geom_sketch_text(aes(label = units), family = "Caveat", nudge_y = 2.5, size = 6) +
  labs(title = "Registered font", x = NULL) +
  theme_sketch(base_family = "Caveat")
```

`ggsketch_check_fonts()` and `register_sketch_font()` need the optional
`systemfonts` package; without it (or without any handwriting font), everything
still renders with the device default — ggsketch never makes fonts a hard
dependency.
