---
title: "Theming & fonts"
description: >
  theme_sketch() light and dark presets, the paper-and-ink palette, and optional
  handwriting fonts.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>", message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 4.2, dpi = 120, dev = "ragg_png",
  out.width = "100%", fig.align = "center"
)
library(ggplot2)
library(ggsketch)
options(ggsketch.seed = 1)
```

The sketchiness of the *marks* comes from the geoms. `theme_sketch()` styles the
surrounding frame — typography, gridlines, background — with a paper-and-ink
palette that complements them.

## Light and dark presets

```{r light}
sales <- data.frame(product = c("Alpha", "Bravo", "Charlie", "Delta"),
                    units   = c(34, 51, 22, 47))

ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Light (paper) preset", x = NULL) +
  theme_sketch()
```

```{r dark}
ggplot(sales, aes(product, units, fill = product)) +
  geom_sketch_col(colour = "grey85", seed = 1L, show.legend = FALSE) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Dark (chalkboard) preset", x = NULL) +
  theme_sketch(dark = TRUE)
```

On a dark background, give the geoms a light outline `colour` (e.g. `"grey85"`)
so the rough strokes read clearly.

## Base size and the rest of the grammar

`theme_sketch()` is a normal ggplot2 theme — combine and override freely.

```{r size}
ggplot(mtcars, aes(wt, mpg)) +
  geom_sketch_point(size = 3, seed = 1L) +
  labs(title = "Bigger base text", subtitle = "theme_sketch(base_size = 15)") +
  theme_sketch(base_size = 15) +
  theme(panel.grid.minor = element_blank())
```

## Handwriting fonts (optional)

The look does **not** depend on fonts — but a handwriting face for the text
completes the effect. Pass a family name to `base_family`, or use `"auto"` to
pick the first installed handwriting font (falling back to the device default if
none are found).

```{r auto, eval = FALSE}
ggplot(sales, aes(product, units)) +
  geom_sketch_col(fill = "#7BAFD4", seed = 1L) +
  theme_sketch(base_family = "auto")
```

Check what is available on your machine:

```{r check-fonts, eval = FALSE}
ggsketch_check_fonts()
#> Available handwriting fonts:
#>   Caveat
```

`ggsketch_check_fonts()` needs the optional `systemfonts` package; without it (or
without any handwriting font), everything still renders with the device default —
ggsketch never makes fonts a hard dependency.
